---
title: "Matrix Check"
author: "Jason D. Everett"
date: "18/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plot Matrix Biomass
```{r, message=FALSE, warning=FALSE}
source("~/GitHub/ZooMSS/fZooMSS_Xtras.R")
suppressPackageStartupMessages({
  library(tidyverse)
  library(patchwork)
  })

enviro <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/enviro_CMIP_Matrix_wPhyto.RDS")
mdl <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_OutofOrder/Output/model_Control.RDS")
Zoo <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_OutofOrder/Output/res_Control.RDS") %>%
  fZooMSS_SpeciesBiomass(mdl) %>%
  fZooMSS_Convert2Tibble(mdl) %>%
  fZooMSS_AddEnviro(enviro) %>%
  pivot_longer(cols = Flagellates:Fish_Large, names_to = "Species", values_to = "Biomass") %>%
  mutate(Chl_log10 = round(Chl_log10,2),
         SST = round(SST,2))

ZooControl <- Zoo

```

```{r, out.width = '100%'}
plot_list <- list()
sp <- mdl$param$Groups$Species

for (i in 1:length(sp)){
  plot_list[[i]] <- Zoo %>%
    filter(Species == sp[i]) %>%
    ggplot(aes(x = SST, y = Chl_log10, colour = Biomass, fill = Biomass)) +
    geom_tile() +
    scale_fill_distiller(palette = "Spectral") +
    scale_colour_distiller(palette = "Spectral") +
    guides(colour = FALSE) + 
    theme_bw() + 
    theme(text = element_text(size = 6), legend.key.width = unit(2,"mm")) + 
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) + 
    ggtitle(sp[i])
}

wrap_plots(plot_list, ncol = 3)

```

There seems to be a lot of inconsistencies in the plot above. The biomass seemingly goes up and down at high frequency at points. Why?

Here is a larger example.

```{r, message=FALSE, warning=FALSE}
plot_list[[4]]
```

Check what is happening with the other experiments (No Omnivores, No Carnivores, No Filter Feeders)
```{r out.width = '100%'}

enviro <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/enviro_CMIP_Matrix_wPhyto.RDS")

mdl <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/NoOmnivores/Output/model_NoOmnivores.RDS")
Zoo <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/NoOmnivores/Output/res_NoOmnivores.RDS") %>%
  fZooMSS_SpeciesBiomass(mdl) %>%
  fZooMSS_Convert2Tibble(mdl) %>%
  fZooMSS_AddEnviro(enviro) %>%
  pivot_longer(cols = Flagellates:Fish_Large, names_to = "Species", values_to = "Biomass") %>%
  mutate(Chl_log10 = round(Chl_log10,2),
         SST = round(SST,2))

plot_list <- list()
sp <- mdl$param$Groups$Species

for (i in 1:length(sp)){
  plot_list[[i]] <- Zoo %>%
    filter(Species == sp[i]) %>%
    ggplot(aes(x = SST, y = Chl_log10, colour = Biomass, fill = Biomass)) +
    geom_tile() +
    scale_fill_distiller(palette = "Spectral") +
    scale_colour_distiller(palette = "Spectral") +
    guides(colour = FALSE) + 
    theme_bw() + 
    theme(text = element_text(size = 6), legend.key.width = unit(2,"mm")) + 
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) + 
    ggtitle(sp[i])
}

wrap_plots(plot_list, ncol = 3)

```

```{r out.width = '100%'}

enviro <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/enviro_CMIP_Matrix_wPhyto.RDS")
mdl <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/NoCarnivores/Output/model_NoCarnivores.RDS")
Zoo <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/NoCarnivores/Output/res_NoCarnivores.RDS") %>%
  fZooMSS_SpeciesBiomass(mdl) %>%
  fZooMSS_Convert2Tibble(mdl) %>%
  fZooMSS_AddEnviro(enviro) %>%
  pivot_longer(cols = Flagellates:Fish_Large, names_to = "Species", values_to = "Biomass") %>%
  mutate(Chl_log10 = round(Chl_log10,2),
         SST = round(SST,2))

plot_list <- list()
sp <- mdl$param$Groups$Species

for (i in 1:length(sp)){
  plot_list[[i]] <- Zoo %>%
    filter(Species == sp[i]) %>%
    ggplot(aes(x = SST, y = Chl_log10, colour = Biomass, fill = Biomass)) +
    geom_tile() +
    scale_fill_distiller(palette = "Spectral") +
    scale_colour_distiller(palette = "Spectral") +
    guides(colour = FALSE) + 
    theme_bw() + 
    theme(text = element_text(size = 6), legend.key.width = unit(2,"mm")) + 
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) + 
    ggtitle(sp[i])
}

wrap_plots(plot_list, ncol = 3)

```



```{r out.width = '100%'}

enviro <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/enviro_CMIP_Matrix_wPhyto.RDS")

mdl <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/NoFilterFeeders/Output/model_NoFilterFeeders.RDS")
Zoo <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/NoFilterFeeders/Output/res_NoFilterFeeders.RDS") %>%
  fZooMSS_SpeciesBiomass(mdl) %>%
  fZooMSS_Convert2Tibble(mdl) %>%
  fZooMSS_AddEnviro(enviro) %>%
  pivot_longer(cols = Flagellates:Fish_Large, names_to = "Species", values_to = "Biomass") %>%
  mutate(Chl_log10 = round(Chl_log10,2),
         SST = round(SST,2))

plot_list <- list()
sp <- mdl$param$Groups$Species

for (i in 1:length(sp)){
  plot_list[[i]] <- Zoo %>%
    filter(Species == sp[i]) %>%
    ggplot(aes(x = SST, y = Chl_log10, colour = Biomass, fill = Biomass)) +
    geom_tile() +
    scale_fill_distiller(palette = "Spectral") +
    scale_colour_distiller(palette = "Spectral") +
    guides(colour = FALSE) + 
    theme_bw() + 
    theme(text = element_text(size = 6), legend.key.width = unit(2,"mm")) + 
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) + 
    ggtitle(sp[i])
}

wrap_plots(plot_list, ncol = 3)

```


These look ok. Lets go back to the control runs and have a look at a set chlorophyll value

```{r, message=FALSE, warning=FALSE}

gg1 <- ZooControl %>%
  filter(Species == sp[4] & Chl_log10 == -1) %>%
  ggplot(aes(x = SST, y = Biomass)) +
  geom_line() + 
  ggtitle(paste0(sp[4],", Chl = 0.1"))

gg2 <- ZooControl %>%
  filter(Species == sp[4] & Chl_log10 == 0) %>%
  ggplot(aes(x = SST, y = Biomass)) +
  geom_line() + 
  ggtitle(paste0(sp[4],", Chl = 1"))

gg3 <- ZooControl %>%
  filter(Species == sp[4] & Chl_log10 == 1) %>%
  ggplot(aes(x = SST, y = Biomass)) +
  geom_line() + 
  ggtitle(paste0(sp[4],", Chl = 10"))

gg1/gg2/gg3

```

There is obviously a fair bit of variability in the output, but there shouldn't be. These changes should be continuous and in the same direction. I wonder if this is related to the oscillations in the time-series data and the way we are averaging the last 50 %?

Lets check the averaging patterns for a single file and see how much difference it makes.

```{r, message=FALSE, warning=FALSE}
enviro_data <- data.frame(chlo = 10, sst = 20)

out <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_Check/RawOutput/20200917_TheMatrix_Control_000001.RDS")

# Function to calculate the mean of the last pd % of the model
fZooMSS_AveOutput2 <- function(out, pd){
  res <- colMeans(out$model$N[(ceiling(dim(out$model$N)[1] - pd*(dim(out$model$N)[1])):dim(out$model$N)[1]),,], dims = 1)
  res2 <- list(res) %>% 
    fZooMSS_SpeciesBiomass(out$model) %>%
    fZooMSS_Convert2Tibble(out$model)
  return(res2)
}

pd <- seq(0.1,0.9, 0.05)
for (p in 1:length(pd)){
  if (p == 1){
    out2 <- fZooMSS_AveOutput2(out, pd[p])
    # out2$pd <- pd[p]
  }
  
  if (p != 1){
    out2[p,] <- fZooMSS_AveOutput2(out, pd[p])
    # out2$pd[p] <- pd[p]
  }
  
}

out2$pd <- pd  

out3 <- out2 %>% 
  pivot_longer(cols = Flagellates:Fish_Large, names_to = "Species", values_to = "Biomass") 

ggplot(data = out3, aes(x = pd, y = Biomass, colour = Species)) + 
  geom_line()
  
```


I don't think the problem is the oscillations and the proportion of the output that we save....

Now lets check that the new run we just did, matches what it shows in the Matrix.

```{r, message=FALSE, warning=FALSE}
# New run
enviro_data <- data.frame(chlo = 10, sst = 20)
out_new <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_Check/RawOutput/20200917_TheMatrix_Control_000001.RDS")
Zoo_new <- list(out_new$abundances) %>% 
  fZooMSS_SpeciesBiomass(out_new$model) %>%
  fZooMSS_Convert2Tibble(out_new$model)


# Find equivalent run in the data
enviro_old <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/enviro_CMIP_Matrix_wPhyto.RDS") %>% 
  filter(chlo == 10, sst == 20)
out_old <- read_rds(paste0("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_OutofOrder/RawOutput/20200917_TheMatrix_Control_",sprintf("%06d",enviro_old$cellID),".RDS"))
Zoo_old <- list(out_old$abundances) %>% 
  fZooMSS_SpeciesBiomass(out_old$model) %>%
  fZooMSS_Convert2Tibble(out_old$model)

Zoo_old
Zoo_new

```

Those runs are different when they should be the same. This suggest that there was a problem with the original runs......

Lets check what the matrix runs look like when I only include the ones where the cellID matches matches what is saved with the file.....

```{r, message=FALSE, warning=FALSE}

enviro <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/enviro_CMIP_Matrix_wPhyto.RDS")
fil <- list.files("/Users/jason/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_OutofOrder/RawOutput/", full.names = TRUE)
enviro$chk_fileID <- NA
enviro$chk_cellID <- NA
enviro$chk_chlo <- NA
enviro$chk_sst <- NA
for (i in 1:length(fil)){
  out <- read_rds(fil[i])
  enviro$chk_fileID[i] = as.numeric(str_sub(fil[i], -10,-5))
  enviro$chk_cellID[i] = out$model$param$cellID
  enviro$chk_chlo[i] = out$model$param$chlo
  enviro$chk_sst[i] = out$model$param$sst
  rm(out)
}

mdl <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_OutofOrder/Output/model_Control.RDS")
Zoo <- read_rds("~/Nextcloud/MME2Work/ZooMSS/_LatestModel/20200917_CMIP_Matrix/Control_OutofOrder/Output/res_Control.RDS") %>%
  fZooMSS_SpeciesBiomass(mdl) %>%
  fZooMSS_Convert2Tibble(mdl) %>%
  fZooMSS_AddEnviro(enviro)


Zoo2 <- Zoo  %>% 
  filter(cellID==enviro$chk_cellID)

enviro2 <- enviro %>% 
  filter(cellID != chk_cellID)

Zoo2 <- Zoo2 %>% 
  pivot_longer(cols = Flagellates:Fish_Large, names_to = "Species", values_to = "Biomass") %>%
  mutate(Chl_log10 = round(Chl_log10,2),
         SST = round(SST,2))

plot_list <- list()
sp <- mdl$param$Groups$Species

for (i in 1:length(sp)){
  plot_list[[i]] <- Zoo2 %>%
    filter(Species == sp[i]) %>%
    ggplot(aes(x = SST, y = Chl_log10, colour = Biomass, fill = Biomass)) +
    geom_tile() +
    scale_fill_distiller(palette = "Spectral") +
    scale_colour_distiller(palette = "Spectral") +
    guides(colour = FALSE) + 
    theme_bw() + 
    theme(text = element_text(size = 6), legend.key.width = unit(2,"mm")) + 
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) + 
    ggtitle(sp[i])
}

wrap_plots(plot_list, ncol = 3)

```

There is no problem here. It looks like there is a problem with the way it ran older simulations in Control. I need to rerun these.

There is also a problem where there is a step-change in the output when Chl > 10. Perhaps this is related to the phytoplankton parameters? Or the above problem

```{r, message=FALSE, warning=FALSE}
gg1 <- ggplot(data = enviro, aes(x = chlo, y = phyto_slope)) +
  geom_line() +
  ggtitle("Phytoplankton Slope")

gg2 <- ggplot(data = enviro, aes(x = chlo, y = phyto_int)) +
  geom_line() +
  ggtitle("Phytoplankton Intercept")

gg3 <- ggplot(data = enviro, aes(x = chlo, y = phyto_max)) +
  geom_line() +
  ggtitle("Phytoplankton Max Size")

gg1/gg2/gg3

```

Doesn't seem to be a problem with the phyto parameters. Wait and see if rerunning the Control runs fixes the problem.